Reliable feature point matching without the use of geometric constraints

### 2013-06-07 text/clustering.tex
V. 2.4.

While the louvain clustering algorithm doesn't require any parameters in 
itself, it tends towards clustering all feature points together in the 
same partition if the graph is very connected.  To ensure that the graph 
is well clustered, the adjacency matrix is pruned so only edges above a 
certain threshold is kept. From empirical analysis keeping around 2.5\% 
of edges seems to work well. Figure \ref{fig:graph} shows the result of 
clustering the feature points displayed as a graph.
%
\begin{SCfigure*}
	\makebox[0.7\textwidth][c]{%
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/MMC_graph_full}
			\caption{Full Graph}
			\label{fig:full_graph}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.3\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/MMC_graph_cropped}
			\caption{Cropped Section of Graph}
			\label{fig:cropped_graph}
		\end{subfigure}%
	}%
	\caption{The partitioned feature graph. Each vertex represents a 
		feature point while an edge represent a high similarity between 
		points. A partition is a connected group marked by a color.  Due 
		to a limited amount of colors some disconnected partitions have 
		been assigned the same color. The edge color of each node 
		signifies which image it belongs to. In the detailed crop to the 
		right the various cluster sizes can be seen, ranging from 
	hundreds of feature points to only two or three.}
	\label{fig:graph}
\end{SCfigure*}

@@

The matching for \emph{MMC} \emph{getMatches} finds matches within all 
partitions with more than two elements as done in \emph{MM}. However as 
can be seen in the example in figure \ref{fig:cropped_graph} many of the 
partitions contain only two feature points from different images linked 
by one edge. In this case we compare the ratio of this match to the next 
best match across all partitions.
%\begin{algorithm}
%\caption{Impl. of getMatches (\emph{from MMC algorithm})}
%\label{alg-getmatches}
%\begin{algorithmic}
%\Require $p$ : set of features, $t\in \mathbb{R}$, $features$ : Set of 
%all features
%\State $matches \gets \varnothing$
%\State $edges \gets \left\{getSimilarity(f_i, f_j) \mid getImg(f_i) \neq 
%getImg(f_j) \wedge f_i, f_j \in p \right\}$
%\If{$\left\vert edges \right\vert = 1$} \Comment If $\exists$ one edge 
%between images
%	\State $f_i,f_j \gets getInterImageFeatures(edges_{inter}, p)$
%	\State $f_m,f_n \gets getTwoNearestNeighbors(f_i, features ~ 
%\backslash ~ \left\{f_i\right\})$
%	\State $f_s,f_t \gets getTwoNearestNeighbors(f_j, features ~ 
%\backslash ~ \left\{f_j\right\})$
%	\State $ratio_i \gets getSimilarity(f_i, f_n) / getSimilarity(f_i, 
%f_j)$
%	\State $ratio_j \gets getSimilarity(f_j, f_t) / getSimilarity(f_j, 
%f_i)$
%	\If{$ratio_i < t\wedge ratio_j < t$}
%		\State $matches \gets matches \cup (f_i, f_j)$
%	\EndIf
%\ElsIf{$\left\vert edges \right\vert > 1$} \Comment If $\exists$ more 
%than one edge between images
%	\ForAll{$f_i \in p$}
%		\State $f_m,f_n \gets getTwoNearestNeighbors(f_i, p ~ \backslash 
%~ \left\{f_i\right\})$
%		\State $ratio \gets getSimilarity(f_i, f_n) / getSimilarity(f_i, 
%f_m)$
%		\If{$getImg(f_i) \neq getImg(f_m) \wedge ratio < t
%\wedge (f_m, f_i) \not\in matches$}
%			\State $matches \gets matches \cup (f_i, f_m)$
%		\EndIf
%	\EndFor
%\EndIf
%
%\Return matches
%\end{algorithmic}
%\end{algorithm}
%
\section{Experiments}
\label{experiment}
%
To reliably measure the accuracy of a matching method on real images we 
either need a set of image pairs tied by a homography or manually count 
the number of inliers. The latter becomes unfeasible the moment we 
attempt to test a non-trivial set of images.

@@

\begin{figure}
	\makebox[0.5\textwidth][c]{%
		\begin{subfigure}[t]{.13\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/graf}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{.27\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/result_graf}
			%\caption{Performance on Scharf}
		\end{subfigure}%
	}%
	\label{fig:results_graf}
	\caption{Accuracy on the \emph{Graf} set from 
		\cite{mikolajczyk2005performance} using img1 and img3. The plot 
	shows the result of $100$ image pairs generated from the source 
images shown to the left. The number of matches is the total number of 
matches for all $100$ image pairs.}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{images/result_viewport}
	\caption{Results on viewport changes using the \emph{Graf} set from 
		\cite{mikolajczyk2005performance} as image pair sources. S is a 
		testst of $100$ crops from img1 and img2. M is likewise but 
	based on img1 and img3.  L is img1 and img4 and XL is img1 and img5.  
To decrease clutter \emph{MM} is not shown, but it performs alike to 
\emph{MMC}}
	\label{fig:result_viewport}
\end{figure}

Figure \ref{fig:result_graf} shows the results for 100 image pairs 
generated from img1 and img3 of the \emph{Graf} image set. The plot 
illustrates how \emph{Isodata} performs well to begin with, but as we 
permit more matches \emph{MM} and \emph{MMC} show superior results. To 
generalize this result, figure \ref{fig:result_accumulated} displays the 
accumulated result over the \emph{Murals} dataset as well as the 
\emph{Graf} pair tested in fig. \ref{fig:result_graf}. In total 900 
image pairs were generated from 9 source pairs. This plot confirms the 
accuracy of \emph{MMC} and \emph{MM} over a bigger data set. As can be 
seen in figure \ref{fig:result_viewport}, the results of \emph{MMC} are 
general across different magnitudes of the viewport change.

For an example of a real life use case \ref{fig:result_pitts} shows the 
results on potentially overlapping image pairs generated from a typical 
holiday scene with partial overlap and slight viewport change.

In terms of performance, table \ref{table:running_times} shows the 
running time of the four algorithms over $100$ image pairs of $250$px by 
$250$px. These numbers should be interpreted with the knowledge that 
much of the code behind \emph{MMC} and \emph{Isodata} is implemented in 
python while \emph{MM} and \emph{SIFT} makes use of opencv to execute 
computationally intensive operations in C++ which renders it much 
faster. While \emph{SIFT}, \emph{MM} and \emph{Isodata} can be 
implemented in $O(nlogn)$ where $n$ is the total amount of feature 
points, \emph{MMC} currently has a complexity of $O(n^2)$ due to the 
construction of a similarity matrix of the feature points. It is however 
possible to approximate this in $O(nlogn)$, something that might be 
pursued in future work.
%
\begin{savenotes}
\begin{table}
	\centering
	\small
\begin{tabular}{l*{4}{c}}
	Algorithm: & \emph{SIFT} & \emph{MM} & \emph{MMC} & \emph{Isodata} 
	\\
	\noalign{\smallskip} 
	%
	Running time: & 20.94s & 18.23s & 883.99s & 555.33s \\
\end{tabular}
\caption{Running times as tested on a Intel(R) Core(TM) i5-3550 CPU @ 
3.30GHz with 8Gb memory}
\label{table:running_times}
\end{table}
\end{savenotes}
%

\begin{figure}
	\makebox[0.5\textwidth][c]{%
		\begin{subfigure}[t]{.15\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/pitts}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{.27\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/result_pitts}
			%\caption{Performance on Scharf}
		\end{subfigure}%
	}%
	\label{fig:results_pitts}
	\caption{Accuracy on an image pair from the Gallagher dataset from 
		\cite{gallagher2008} using cropped versions of 100\_1941 and 
		100\_1942. The plot shows the result of $100$ image pairs 
		generated from the source images shown to the left.  The number 
	of matches is the total number of matches for all $100$ image 
pairs.}
\end{figure}

\section{Summary}
We have addressed the problem of matching feature points without using 
geometrical constraints and proposed two algorithms \emph{Mirror Match 
(MM)} and \emph{Mirror Match with Clustering (MMC)}. The two algorithms 
share the same common idea that feature points should have better 
matches in another image than in the image they came from to be 
considered good matches and \emph{MMC} further improves on this idea by 
using the structure of the similarity graph of the feature points. We 
also address how to test the algorithms by constructing the 
\emph{Murals} dataset consisting of $8$ source image pairs and a way to 
generate testsets based on these pairs.  The proposed approach shows 
promising results on both the \emph{Murals} dataset as well as a real 
life example from the Gallagher dataset and the \emph{Graf} image set, 
beating even a method using geometrical constraints when matching image 
pairs that might not overlap.
%
\bibliographystyle{IEEEtran}
\bibliography{bibliography}
\end{document}

>>>>

While the louvain clustering algorithm doesn't require any parameters in 
itself, it tends towards clustering all feature points together in the 
same partition if the graph is very connected.  To ensure that the graph 
is well clustered, the adjacency matrix is pruned so only edges above a 
certain threshold is kept. From empirical analysis keeping around 2.5\% 
of edges seems to work well. Figure \ref{fig:graph} shows the result of 
clustering the feature points displayed as a graph.
%
\begin{SCfigure*}
	\makebox[0.7\textwidth][c]{%
		\begin{subfigure}[t]{0.28\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/MMC_graph_full}
			\caption{Full Graph}
			\label{fig:full_graph}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.28\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/MMC_graph_cropped}
			\caption{Cropped Section of Graph}
			\label{fig:cropped_graph}
		\end{subfigure}%
	}%
	\caption{The partitioned feature graph. Each vertex represents a 
		feature point while an edge represent a high similarity between 
		points. A partition is a connected group marked by a color.  Due 
		to a limited amount of colors some disconnected partitions have 
		been assigned the same color. The edge color of each node 
		signifies which image it belongs to. In the detailed crop to the 
		right the various cluster sizes can be seen, ranging from 
	hundreds of feature points to only two or three.}
	\label{fig:graph}
\end{SCfigure*}

@@

The matching for \emph{MMC} \emph{getMatches} finds matches within all 
partitions with more than two elements as done in \emph{MM}. However as 
can be seen in the example in figure \ref{fig:cropped_graph} many of the 
partitions contain only two feature points from different images linked 
by one edge. In this case we compare the similarity of this match to the 
next best match across all partitions and filter the matches based on 
this ratio like in \emph{SIFT}.
%\begin{algorithm}
%\caption{Impl. of getMatches (\emph{from MMC algorithm})}
%\label{alg-getmatches}
%\begin{algorithmic}
%\Require $p$ : set of features, $t\in \mathbb{R}$, $features$ : Set of 
%all features
%\State $matches \gets \varnothing$
%\State $edges \gets \left\{getSimilarity(f_i, f_j) \mid getImg(f_i) \neq 
%getImg(f_j) \wedge f_i, f_j \in p \right\}$
%\If{$\left\vert edges \right\vert = 1$} \Comment If $\exists$ one edge 
%between images
%	\State $f_i,f_j \gets getInterImageFeatures(edges_{inter}, p)$
%	\State $f_m,f_n \gets getTwoNearestNeighbors(f_i, features ~ 
%\backslash ~ \left\{f_i\right\})$
%	\State $f_s,f_t \gets getTwoNearestNeighbors(f_j, features ~ 
%\backslash ~ \left\{f_j\right\})$
%	\State $ratio_i \gets getSimilarity(f_i, f_n) / getSimilarity(f_i, 
%f_j)$
%	\State $ratio_j \gets getSimilarity(f_j, f_t) / getSimilarity(f_j, 
%f_i)$
%	\If{$ratio_i < t\wedge ratio_j < t$}
%		\State $matches \gets matches \cup (f_i, f_j)$
%	\EndIf
%\ElsIf{$\left\vert edges \right\vert > 1$} \Comment If $\exists$ more 
%than one edge between images
%	\ForAll{$f_i \in p$}
%		\State $f_m,f_n \gets getTwoNearestNeighbors(f_i, p ~ \backslash 
%~ \left\{f_i\right\})$
%		\State $ratio \gets getSimilarity(f_i, f_n) / getSimilarity(f_i, 
%f_m)$
%		\If{$getImg(f_i) \neq getImg(f_m) \wedge ratio < t
%\wedge (f_m, f_i) \not\in matches$}
%			\State $matches \gets matches \cup (f_i, f_m)$
%		\EndIf
%	\EndFor
%\EndIf
%
%\Return matches
%\end{algorithmic}
%\end{algorithm}
%
\section{Experiments}
\label{experiment}
%
To reliably measure the accuracy of a matching method on real images we 
either need a set of image pairs tied by a homography or manually count 
the number of inliers. The latter becomes unfeasible the moment we 
attempt to test a non-trivial set of images.

@@

\begin{figure}
	\makebox[0.5\textwidth][c]{%
		\begin{subfigure}[t]{.13\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/graf}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{.27\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/result_graf}
			%\caption{Performance on Scharf}
		\end{subfigure}%
	}%
	\caption{Accuracy on the \emph{Graf} set from 
		\cite{mikolajczyk2005performance} using img1 and img3. The plot 
		shows the result of $100$ image pairs generated from the source 
		images shown to the left. The number of matches is the total 
		number of matches for all $100$ image pairs.}
	\label{fig:result_graf}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{images/result_viewport}
	\caption{Results on viewport changes using the \emph{Graf} set from 
		\cite{mikolajczyk2005performance} as image pair sources. Img 1 
		to 6 in the \emph{Graph} set are images of the same subject but 
		with gradually increasing view port change. Img1 and 2 are 
		almost identical while img1 and 6 are taken from very different 
		angles. S is a testst of $100$ crops from img1 and img2. M is 
	likewise but based on img1 and img3.  L is img1 and img4 and XL is 
img1 and img5.  To decrease clutter \emph{MM} is not shown, but it 
performs alike to \emph{MMC}}
	\label{fig:result_viewport}
\end{figure}

Figure \ref{fig:result_graf} shows the results for 100 image pairs 
generated from img1 and img3 of the \emph{Graf} image set. The plot 
illustrates the results of \emph{SIFT}, \emph{MM}, \emph{MMC} and 
\emph{Isodata} as thresholds are varied over a range. \emph{Isodata} 
performs well with strict thresholds, but when we require more matches 
\emph{MM} and \emph{MMC} show superior results. To see if this results 
generalizes over more image sets, we tested the four algorithms on the 
\emph{Murals} dataset as well as the \emph{Graf} pair tested in fig.  
\ref{fig:result_graf}. In total 900 image pairs were generated from 9 
source pairs generating the plot shown in figure 
\ref{fig:result_accumulated} displaying the accumulated amount of 
matches on the x-axis. To verify that this performance is independent of 
the magnitude of view port change, we tested \emph{MMC} and \emph{SIFT} 
on the \emph{Graf} image set as shown in figure 
\ref{fig:result_viewport}. The figure confirms that \emph{MMC} is 
generally superior across view port changes.

For an example of a real life use case figure \ref{fig:result_pitts} 
shows the results on potentially overlapping image pairs generated from 
a typical holiday scene with partial overlap and slight view port 
change.

In terms of performance, table \ref{table:running_times} shows the 
running time of the four algorithms over $100$ image pairs of $250$px by 
$250$px. These numbers should be interpreted with the knowledge that 
much of the code behind \emph{MMC} and \emph{Isodata} is implemented in 
python while \emph{MM} and \emph{SIFT} makes use of opencv to execute 
computationally intensive operations in C++ which renders it much 
faster. While \emph{SIFT}, \emph{MM} and \emph{Isodata} can be 
implemented in $O(nlogn)$ where $n$ is the total amount of feature 
points, \emph{MMC} currently has a complexity of $O(n^2)$ due to the 
construction of a similarity matrix of the feature points. It is however 
possible to approximate this in $O(nlogn)$ using search trees, something 
that could be pursued in future work.
%
\begin{savenotes}
\begin{table}
	\centering
	\small
\begin{tabular}{l*{4}{c}}
	Algorithm: & \emph{SIFT} & \emph{MM} & \emph{MMC} & \emph{Isodata} 
	\\
	\noalign{\smallskip} 
	%
	Running time: & 20.94s & 18.23s & 883.99s & 555.33s \\
\end{tabular}
\caption{Running times as tested on a Intel(R) Core(TM) i5-3550 CPU @ 
3.30GHz with 8Gb memory}
\label{table:running_times}
\end{table}
\end{savenotes}
%

\begin{figure}
	\makebox[0.5\textwidth][c]{%
		\begin{subfigure}[t]{.15\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/pitts}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{.27\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/result_pitts}
			%\caption{Performance on Scharf}
		\end{subfigure}%
	}%
	\caption{Accuracy on an image pair from the Gallagher dataset from 
		\cite{gallagher2008} using cropped versions of 100\_1941 and 
		100\_1942. The plot shows the result of $100$ image pairs 
		generated from the source images shown to the left.  The number 
	of matches is the total number of matches for all $100$ image 
pairs.}
	\label{fig:result_pitts}
\end{figure}

\section{Summary}
We have addressed the problem of matching feature points without using 
geometrical constraints and proposed two algorithms \emph{Mirror Match 
(MM)} and \emph{Mirror Match with Clustering (MMC)}. The two algorithms 
share the same common idea that feature points should have better 
matches in another image than in the image they came from to be 
considered good matches and \emph{MMC} further improves on this idea by 
using the structure of the similarity graph of the feature points. We 
also address how to test the algorithms by constructing the 
\emph{Murals} dataset consisting of $8$ source image pairs and a way to 
generate testsets based on these pairs.  The proposed approach shows 
promising results on both the \emph{Murals} dataset as well as a real 
life example from the Gallagher dataset and the \emph{Graf} image set, 
surpassing even a method using geometrical constraints when matching 
image pairs that might not overlap.
%
\bibliographystyle{IEEEtran}
\bibliography{bibliography}
\end{document}
